---
title: "ProblemSet4"
author: "Yangning Tan"
format: html
editor: visual
---

## Problem 1 - Tidyverse

```{r}
# install package
library(nycflights13)
library(tidyverse)
```

1.  Generate a table (which can just be a nicely printed tibble) reporting the mean and median departure delay per airport. Generate a second table (which again can be a nicely printed tibble) reporting the mean and median arrival delay per airport. Exclude any destination with under 10 flights. Do this exclusion through code, not manually.

    Additionally,

    -   Order both tables in descending mean delay.

    -   Both tables should use the airport *names* not the airport *codes*.

    -   Both tables should print all rows.

    ```{r}
    # mean and median of departure delay
    tb_flights <- nycflights13::flights

    tb_flights %>% 
      inner_join(nycflights13::airports, by = c("origin" = "faa")) %>% 
      select(name, dep_delay) %>% 
      group_by(name) %>% 
      summarise(dep_delay_mean = mean(dep_delay, na.rm = TRUE),
                dep_delay_median = median(dep_delay, na.rm = TRUE)) %>% 
      arrange(desc(dep_delay_mean))
    ```

    ```{r}
    # mean and median of arrival delay
    tb_flights %>%
      inner_join(nycflights13::airports, by = c("dest" = "faa")) %>%
      group_by(dest) %>%
      filter(n() >= 10) %>%
      ungroup() %>% 
      select(name, arr_delay) %>%
      group_by(name) %>%
      summarise(arr_delay_mean = mean(arr_delay, na.rm = TRUE),
                arr_delay_median = median(arr_delay, na.rm = TRUE)) %>% 
      arrange(desc(arr_delay_mean))
    ```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
